Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-19 paper.tex
minor fix
sec 4.0

\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) #2}

@@

We captures the set of operators that every cast representation must provide 
with \emph{cast abstract data type} (Definition~\ref{def:cast-rep}). 
The first three operators allow a language implementation to construct casts. 
So we call them \textit{cast constructors}. And the forth and last operator 
enable the implementation to use apply casts.

@@

To use our framework, users needs to prove that their cast representation 
implements Lazy D Cast ADT (Definition~\ref{def:surely-lazyd}). To be a Lazy D 
Cast ADT, $ id(T) $ must acts like the identity function. And $ seq(c,c) $ 
must acts like a sequence of casts. What's more, $ cast(T_1,l,T_2) $ must act 
like the Lazy D $ applyCast $ (Fig.~\ref{fig:applyCast-D-C}).

\begin{definition}[Lazy D Cast ADT]
	\label{def:surely-lazyd}
	A  Cast ADT is a Lazy D Cast ADT if:
	\begin{enumerate}
		\item If $ v : T $, then $ applyCast(v,id(T)) = \mathtt{succ} \; v $
		\item If $ \judgeType{v}{T_1} $,
		$ \judgeTypeFT{c_1}{T_1}{T_2} $, and
		$ \judgeTypeFT{c_2}{T_2}{T_3} $,\\
		then $ applyCast(v,seq(c_1,c_2)) = 
		applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2) $ \\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item If $ v : T_1 $ and $ \neg T_1 \smile T_2 $,
		then $ applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
		\item If $ v : \star $, 
		then $ applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
		\item If $ v : P_1 $,
		then $ applyCast(\hcvOOinj{P_1}{v},cast(\star,l,P_2)) 
		= applyCast(v,cast(P_1,l,P_2)) $
		\item If $ v : P $,
		then $ applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item If $ v : \POOunit $,
		then $ applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
		\item
		$ 
		applyCast(\hcvOOfun{c_1}{\rho}{x}{b}{c_2}, 
		cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$

		\item $ 
		applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
		\times 
		T_4)) $ \\
		$ = 
		\rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
		$ 
		\item $ 
		applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
		\item $ 
		applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
	\end{enumerate}
\end{definition}

@@

If a cast representation is a Lazy D Cast ADT, people can apply our theorems 
and be confident that this cast representation is valid.

Some cast representations are also \textit{monoids}. This definition is useful 
in the proof of our framework, and might be interesting for other people.

\begin{definition}[Monoid]
	A Cast ADT is a monoid if 
	for all
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c_1) = c_1$,
		\item $seq(c_1,id(T_2)) = c_1$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{definition}

In subsection~\ref{ssec:framework:cek}, we introduce \effCEK{C}, a family 
of space-efficient semantics indexed by Cast ADT ($ C $). When $ C $ is a Lazy 
D Cast ADT, \effCEK{C} repects \ineffCEKD.
In subsection~\ref{secc:framework:monoid-correct}, we prove when $ C $ is
a Lazy D Cast ADT and a monoid as well, $ \judgeDeval{e}{o} $ if 
and only if $ \judgeSeval{C}{e}{o} $. In 
subsection~\ref{secc:framework:all-correct}, we further prove that when $ C $ is
a Lazy D Cast ADT, \evalEqv{\ineffCEKD}{\effCEK{C}}.

>>>>

\newcommand{\judgeDeval}[2]{eval_{\mathcal{D}}(#1) = #2}

@@

We captures the set of operators that every cast representation must provide 
with \emph{cast abstract data type} (Definition~\ref{def:cast-rep}). 
The first three operators allow a language implementation to construct casts. 
So we call them \textit{cast constructors}. And the forth and last operator 
enable the implementation to apply casts on values.

@@

Some instances of Cast ADT (or "some Cast" for short) are also 
\textit{monoids}. This definition is useful in the proof of 
our framework, and might be interesting for other people.

\begin{definition}[Monoid]
	A Cast is a monoid if 
	for all
	$c_1 : T_1 \Longrightarrow T_2$,
	$c_2 : T_2 \Longrightarrow T_3$, and
	$c_3 : T_3 \Longrightarrow T_4$,
	\begin{enumerate}
		\item $seq(id(T_1),c_1) = c_1$,
		\item $seq(c_1,id(T_2)) = c_1$, and
		\item $seq(seq(c_1, c_2), c_3) = seq(c_1, seq(c_2, c_3))$.
	\end{enumerate}
\end{definition}

To use our framework, users needs to prove that their cast representation
implements Lazy D Cast ADT (Definition~\ref{def:surely-lazyd}). Property (1) 
states that $ id(T) $ must acts like the identity function. Property (2) states 
that $ seq(c,c) $ must acts like a sequence of casts. Properties (3) through 
(11) state that $ cast(T_1,l,T_2) $ must act like the Lazy D $ applyCast $ 
defined in Fig.~\ref{fig:applyCast-D-C}.

\begin{definition}[Lazy D Cast ADT]
	\label{def:surely-lazyd}
	A  Cast is a Lazy D Cast if:
	\begin{enumerate}
		\item If $ v : T $, then $ applyCast(v,id(T)) = \mathtt{succ} \; v $
		\item If $ \judgeType{v}{T_1} $,
		$ \judgeTypeFT{c_1}{T_1}{T_2} $, and
		$ \judgeTypeFT{c_2}{T_2}{T_3} $,\\
		then $ applyCast(v,seq(c_1,c_2)) = 
		applyCast(v,c_1) >>= \lambda v.applyCast(v,c_2) $ \\
		where 
		\[
		\begin{array}{rcl}
		\rOOsucc{v} >>= f & = & f(v) \\
		\rOOfail{l} >>= f & = & \rOOfail{l}
		\end{array}
		\]
		\item If $ v : T_1 $ and $ \neg T_1 \smile T_2 $,
		then $ applyCast(v,cast(T_1, l, T_2)) = \rOOfail{l} $
		\item If $ v : \star $, 
		then $ applyCast(v,cast(\TOOdyn,l,\TOOdyn)) = \rOOsucc{v} $
		\item If $ v : P_1 $,
		then $ applyCast(\hcvOOinj{P_1}{v},cast(\star,l,P_2)) 
		= applyCast(v,cast(P_1,l,P_2)) $
		\item If $ v : P $,
		then $ applyCast(v,cast(P,l,\star)) = \rOOsucc{(\hcvOOinj{P}{v})} $
		\item If $ v : \POOunit $,
		then $ applyCast(v,cast(\POOunit,l,\POOunit)) = \rOOsucc{v} $
		\item If $ (\hcvOOfun{c_1}{\rho}{x}{e}{c_2}) : \POOfun{T_1}{T_2} $,
		then\\
		$ 
		applyCast(\hcvOOfun{c_1}{\rho}{x}{e}{c_2}, 
		cast(\POOfun{T_1}{T_2},l,\POOfun{T_3}{T_4})) \\
		= 
		\rOOsucc{(\hcvOOfun{seq(cast(T_3,l,T_1),c_1)}{\rho}{x}{b}{seq(c_2,cast(T_2,l,T_4))})}$
		\item If $ (\hcvOOcons{v_1}{c_1}{v_2}{c_2}) : \POOprod{T_1}{T_2} $,
		then \\
		$ 
		applyCast(\hcvOOcons{v_1}{c_1}{v_2}{c_2},cast(\POOprod{T_1}{T_2},l,T_3 
		\times 
		T_4)) $ \\
		$ = 
		\rOOsucc{(\hcvOOcons{v_1}{seq(c_1,cast(T_1,l,T_3))}{v_2}{seq(c_2,cast(T_2,l,T_4))})}
		$ 
		\item If $ (\hcvOOinl{v}{c}) : \POOsum{T_1}{T_2} $ ,
		then \\
		$ 
		applyCast(\hcvOOinl{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinl{v}{seq(c,cast(T_1,l,T_3))})} $
		\item If $ (\hcvOOinr{v}{c}) : \POOsum{T_1}{T_2} $ ,
		then \\ $ 
		applyCast(\hcvOOinr{v}{c},cast(\POOsum{T_1}{T_2},l,\POOsum{T_3}{T_4}))
		= \rOOsucc{(\hcvOOinr{v}{seq(c,cast(T_2,l,T_4))})} $
	\end{enumerate}
\end{definition}

@@

After showing their cast representation is a Lazy D Cast, users of our 
frameworks can plug-in this cast representation to our space-efficient abstract 
machine in subsection~\ref{ssec:framework:cek}, and prove that this machine is 
equivalent to \ineffCEKD subsection~\ref{secc:framework:all-correct}, which 
depends on lemmas in subsection~\ref{secc:framework:monoid-correct}.
